### 2013-12-17 msr/latex/paper1.tex
tarted working on reviews from Peter.

\documentclass{acm_proc_article-sp}

\begin{document}
\title{On Reallocation of Resources during Releases \titlenote{Permission to make digital or hard copies of all or part of this work for
personal or classrooms use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. To copy, to republish, to post on servers or to redistribute to lists,
requires prior specific permission and/or a fee. Copyright \copyright 2013}}
\numberofauthors{2}
\author{
% 1st. author
\alignauthor
Md Tajmilur Rahman\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \email{mdt\_rahm@encs.concordia.ca}
% 2nd. author
\alignauthor
Peter C. Rigby\\
       \affaddr{Concordia University}\\
       \affaddr{Montreal, QC H3G 1M8}\\
       \email{peter.rigby@concordia.ca}
}
\date{20 Oct 2013}
\maketitle

\begin{abstract}
Different software projects practices different release models to compete in the field of software development. From the historical data of a large project we quantify aspects of developers' distribution, code ownership, different release periods, focus on the areas of the code-base, percentage of working with owned files during different periods within a release cycle for this project. This study reveals a unique process performing well extraction of data from Git for the development of a large open source project by which we can perceive well understanding on the release process and distribution of developers throughout the release providing us significant points of information that can lead us to further decisional analysis aiming novel solutions to the disruptive events for the practitioners of software development.
\end{abstract}

\category{K.6.3}{Software Management}{Software Development, Software Resource Management, Resource Reallocation}
\terms{Experiment, Human Factors, Resource Management, Reallocation}
\keywords{Resource Reallocation, Code Ownership, Software Releases} % NOT required for Proceedings

\end{document}